---
title: "Uncertainty of signal detection measures"
author: "Monica Thieu"
date: "5/17/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(tidyverse)
require(rlang)
source("../R/utils-sdt.R")

super_spread <- function (data, key, ..., sep = "_", fill = NA, convert = FALSE, drop = TRUE) {
  dots <- exprs(...)
  key <- enquo(key)
  output <- data %>%
    gather(gkey, value, !!! dots) %>%
    unite(ukey, !! key, gkey, sep = sep) %>%
    spread(ukey, value, fill = fill, convert = convert, drop = drop)
  
  return (output)
}
```

Signal detection theory provides a framework for determining the ability of an observer to discriminate "signal" from "noise".

It stands to reason that for a given observer, with some underlying sensitivity, their performance on a signal detection task would not be the same every time they did the task, but would take the form of a distribution centered on their most likely performance. Thus, our goal in administering a single signal detection task to an observer is to estimate the center of that distribution, to find their most likely performance.

It makes sense that administering more trials to an observer would yield a less noisy estimate of their performance. Signal detection measures, however, crucially _do not have an associated standard error that depends on the number of trials completed by an observer._ 

```{r}
# ASSUME equal number of signal and noise trials presented

n_iterations <- 1000

params <- crossing(this_dprime = c(0.5, 1:3),
                   this_c = c(-0.5, 0, 0.5),
                   n_trials_per_iteration = c(20, 100, 200)) %>%
  group_by(this_dprime, this_c, n_trials_per_iteration) %>%
  do(sim = data.frame(type = rep(c("signal", "noise"), times = .$n_trials_per_iteration/2),
                      param_dprime = .$this_dprime,
                      param_c = .$this_c) %>%
       slice(rep(1:n(), times = n_iterations)) %>%
       mutate(iteration = rep(1:n_iterations, each = n() / n_iterations)) %>%
       group_by(iteration, type) %>%
       mutate(n_trials = n(),
              latent = if_else(type == "noise",
                               rnorm(n(), -param_dprime/2, 1), # centering d' so c = 0 is unbiased
                               rnorm(n(), param_dprime/2, 1)),
              resp = if_else(latent >= param_c,
                             "signal",
                             "noise")) %>%
       group_by(iteration, param_dprime, param_c, n_trials, type, resp) %>%
       summarize(rate = n() / unique(n_trials),
                 rate_snodgrass = (n() + .5) / (unique(n_trials) + 1)) %>% # snodgrass correction
       complete(iteration, type, resp) %>%
       distinct() %>% # not sure why but complete() is creating duplicate rows
       group_by(iteration, param_dprime, param_c, n_trials) %>%
       mutate(rate = coalesce(rate, 0),
              rate_snodgrass = coalesce(rate_snodgrass, .5 / (unique(n_trials) + 1))) %>% # need to populate 0 false alarms row in case of perfect performance
       filter(resp == "signal") %>%
       select(-resp) %>%
       mutate(type = recode(type,
                            signal = "hit",
                            noise = "fa")) %>%
       super_spread(type, starts_with("rate")) %>%
       summarize(dprime = sdt_dprime(hit_rate_snodgrass, fa_rate_snodgrass),
                 sdt_c = sdt_c(hit_rate_snodgrass, fa_rate_snodgrass),
                 aprime = sdt_aprime(hit_rate_snodgrass, fa_rate_snodgrass),
                 b2prime = sdt_b2prime(hit_rate_snodgrass, fa_rate_snodgrass)))

sims <- bind_rows(params$sim)

sims_summary <- sims %>%
  group_by(param_dprime, param_c, n_trials) %>%
  summarize_at(vars(dprime, sdt_c, aprime, b2prime), .funs = c("median", "mad"))
```

```{r}
sims %>%
  ggplot(aes(x = dprime, color = factor(n_trials), fill = factor(n_trials))) +
  #geom_histogram(bins = 20, position = "dodge", size = 0.2, alpha = 0.3) +
  geom_freqpoly(bins = 20) +
  geom_vline(aes(xintercept = dprime_median, color = factor(n_trials)), data = sims_summary) +
  geom_vline(aes(xintercept = param_dprime), linetype = 3) +
  facet_grid(param_c ~ param_dprime) +
  theme_bw()
```

```{r}
sims %>%
  ggplot(aes(x = sdt_c, color = factor(n_trials), fill = factor(n_trials))) +
  #geom_histogram(bins = 20, position = "dodge", size = 0.2, alpha = 0.3) +
  geom_freqpoly(bins = 20) +
  geom_vline(aes(xintercept = sdt_c_median, color = factor(n_trials)), data = sims_summary) +
  geom_vline(aes(xintercept = param_c), linetype = 3) +
  facet_grid(param_c ~ param_dprime) +
  theme_bw()
```

```{r}
sims %>%
  ggplot(aes(x = aprime, color = factor(n_trials), fill = factor(n_trials))) +
  #geom_histogram(bins = 20, position = "dodge", size = 0.2, alpha = 0.3) +
  geom_freqpoly(bins = 20) +
  geom_vline(aes(xintercept = aprime_median, color = factor(n_trials)), data = sims_summary) +
  facet_grid(param_c ~ param_dprime, scales = "free_x") +
  theme_bw()
```